---
title: "STAT 222 Spring 2022 HW10"
author: "Matthew Zhao"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin = .65in
fontsize: 12pt
---

```{r include=F}
library(ggplot2)
library(knitr)
library(emmeans)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F, collapse=T)
options(scipen=6, digits=7)
# options(show.signif.stars = FALSE) # Show no stars for significance tests
```



# Question 1 --- Auditor Training
```{r}
auditor = read.table("http://www.stat.uchicago.edu/~yibi/s222/AuditorTraining", header=T)
```

## Q1a --- 2 points

```{r}
auditor$treatment = as.factor(auditor$treatment)
auditor$block = as.factor(auditor$block)

with(auditor, interaction.plot(block,treatment,y,type='b'))
# ggplot(data = auditor, mapping = aes(x=block, y=y,color = treatment)) + 
#   geom_line() + geom_point() + xlab("Block") + ylab("Response")
```


It is not reasonable to assume no treatment-block interaction since the lines are not parallel.

## Q1b --- 3 points

```{r}
lm <- lm(data = auditor, y~treatment+block)
anova(lm)
mean(~y,data=auditor)
sort(mean(y~treatment,data=auditor))
```


$SS_{trt} = \displaystyle b\sum_{i=1}^g(\bar{y}_{i\cdot}-\bar{y}_{\cdot\cdot})^2$

$\displaystyle = 10*((70.6-77.1)^2 + (74.6-77.1)^2 + (86.1-77.1)^2) = 1295$

## Q1c --- 4 points

```{r}
lmemmeans <- emmeans(lm,~treatment)
sprintf("Tukey's HSD: %f",qtukey(1-0.05,3,18)/sqrt(2) * sqrt(6.24*((1/10)+(1/10)))) # hsd
summary(contrast(lmemmeans,method="pairwise",adjust="tukey"),infer=c(T,T), level=0.95)
```

## Q1d --- 4 points

```{r}
lm1 <- lm(data = auditor,y~treatment)
anova(lm1)
```

```{r}
sprintf("New Tukey's HSD: %f",qtukey(1-0.05,3,27)/sqrt(2) * sqrt(20.21*((1/10)+(1/10)))) # hsd
lm1emmeans <- emmeans(lm1,~treatment)
summary(contrast(lm1emmeans,method="pairwise",adjust="tukey"),infer=c(T,T), level=0.95)
```


Only the latter two pairs are significant (1-3),(2-3).

# Question 2

```{r}
mydata = read.table("http://users.stat.umn.edu/~gary/book/fcdae.data/pr13.4", header=T)
```

## Q2a --- 2 points

It is a Latin Squares design with two squares reusing columns (grader block).

## Q2b --- 2 points

```{r}
lm2 <- lm(score~as.factor(exam)+as.factor(student)+as.factor(grader),data = mydata)
anova(lm2)
```


No because the p value for the exams treatment is significantly smaller than 0.05,0.01,etc. This means that we can reject the null that the effect of different exams on score is zero, and so we conclude that exams are not equivalent in the level difficulty since exams have an effect on score.

## Q2c --- 4 points

```{r}
lm2emmeans <- emmeans(lm2,~exam)
summary(contrast(lm2emmeans,method=list(c_1 = c(-1,0,1/3,1/3,1/3),c_2 = c(0,-1,1/3,1/3,1/3)),adjust="bonferroni"),infer=c(F,F), level=0.95)
bonf <- qt(0.05/2/2,32,lower.tail = F)
se <- sqrt(13.91 * ((3*(((1/3)^2)/10)) + (1/10))) #se
bonf*se
```

```{r}
sprintf("CI C1: (%f, %f)",8.1-bonf*se,8.1+bonf*se)
sprintf("CI C2: (%f, %f)",-9.8-bonf*se,-9.8+bonf*se)
```


From contrast 1, since the 95% CI only contains positive values, we can conclude that exam A is more difficult than the old exams with previous exams having higher average scores than exam A. From contrast 2, sinec the 95% CI only contains negative values, we can conclude that exam B is easier than old exams with previous exams having lower average scores than exam B. 

## Q2d --- 4 points

```{r}
sprintf("Tukey's HSD: %f",qtukey(1-0.05,5,32)/sqrt(2) * sqrt(13.91*((1/10)+(1/10)))) # hsd
lm2est <- emmeans(lm2,~grader)
summary(contrast(lm2est,method="pairwise",adjust="tukey"),infer=c(F,F), level=0.95)
```

```
   5      4        3       1      2  
 59.5    64.7    65.4    66.0   68.6 
        -----------------------------  
 -----------------------------
```
