---
title: 'STAT 222 Spring 2022 HW2'
author: "Matthew Zhao"
fontsize: 12pt
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin=0.5in
fig_crop: no
---

```{r include=F}
library(knitr)
library(ggplot2)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F, collapse=TRUE)
options(width=70, digits=5, scipen=8)
options(show.signif.stars = FALSE) # Show no stars for significance tests
```
```{r include=F}
library(mosaic)
```

## Question 1

```{r}
pr3.2 = read.table("http://users.stat.umn.edu/~gary/book/fcdae.data/pr3.2",h=T)
pr3.2$trt = factor(pr3.2$trt, labels=c("none", "1 pregnant", "1 virgin",
                    "8 pregnant", "8 virgin"))
```

```{r, fig.width=3.5, fig.height=2.8, out.width='50%'}
library(ggplot2)
ggplot(pr3.2, aes(x=trt, y=days))+ geom_point() +
  ylab("lifespan of male fruit flies (days)")+xlab("Treatment")
```

### Q1a --- 5 points

```{r, message=FALSE}
library(mosaic)
mean(days ~ trt, data=pr3.2)
```
```{r}
anova(lm(days ~ trt, data = pr3.2))
```
$\text{LSD}= t_{N-g,\frac{\alpha}{2}} * \sqrt{\text{MSE}(\frac{1}{n}+\frac{1}{n})}$
We find the value for $t_{N-g,\frac{\alpha}{2}}$ here:
```{r}
qt(0.05/2,125-5,lower.tail=F)
```
For $\sqrt{\text{MSE}(\frac{1}{n}+\frac{1}{n})}$, each group has $n=25$ observations and the MSE is taken from the ANOVA table above and is equal to 219 Therefore, $\sqrt{219(\frac{1}{25}+\frac{1}{25})}=15.453$ and so LSD $=1.9799*4.1857 = 8.2873$.


Below we calculate the group means and :
```{r echo=F}
fly_means <- pr3.2 %>% group_by(trt) %>% summarise(group_mean = mean(days))
for (x in 1:(nrow(fly_means)-1)) {
  for (y in (x+1):nrow(fly_means)) {
    print(paste0(as.character(fly_means[x,1]),"-",
                 as.character(fly_means[y,1]), " ",
                 round(abs(fly_means$group_mean[x]-
                             fly_means$group_mean[y]),3),
                 "   ",
                 ifelse(abs(fly_means$group_mean[x]-
                              fly_means$group_mean[y]) < 8.2873,
                        "Not Significant", "Significant")))
  }
}
```
Underline Diagram
```
  5 (8v)      4 (8p)       1 (none)      2 (1p)      3 (1v)
   38.72       56.76         63.36        63.56      64.80
            ------------------------------------------------
```

### Q1b --- 6 points

Construct a 95% confidence for each of the contrasts below;

* $C_1 = \mu_0 - (\mu_{8p}+\mu_{8v})/2$ \dotfill (none v.s. 8 female companion)
* $C_2 = (\mu_{1v}+\mu_{8v})/2-(\mu_{1p}+\mu_{8p})/2$ \dotfill (virgin v.s. pregnant)


A 95% confidence interval for a contrast equals $\hat{C}_{k\cdot} \pm t_{N-g,\frac{\alpha}{2}} \times SE(\hat{C})$ where $\displaystyle SE(\hat{C}) = \sqrt{MSE* \sum_{i=1}^{g}\frac{c_{i}^{2}}{n_{i}}}$


For the first contrast:<br>
$\hat{C_{1}} = \displaystyle \sum_{i=1}^{g} c_{i}\bar{y}_{i\cdot} = 1*\bar{y}_{0\cdot} + -0.5*\bar{y}_{8p\cdot}+-0.5*\bar{y}_{8v\cdot} = 1*63.36 + -0.5*56.76 + -0.5*38.72 =15.62$


$\displaystyle SE(\hat{C_{1}}) = \sqrt{MSE* \sum_{i=1}^{g}\frac{c_{i}^{2}}{n_{i}}} = \sqrt{219 * (\frac{1^2}{25}} + \frac{(-0.5)^2}{25}} + \frac{(-0.5)^2}{25})} = 3.6249$


CI: $15.62 \pm 1.9799 \times 3.6249 = C_{1} \in (8.4431,22.797)$


For second contrast:<br>
$\hat{C_{2} = \displaystyle \sum_{i=1}^{g} c_{i}\bar{y}_{i\cdot} = 0.5*\bar{y}_{1v\cdot} + 0.5*\bar{y}_{8v\cdot}+-0.5*\bar{y}_{1p\cdot} +-0.5*\bar{y}_{8p\cdot}= 0.5*64.80 + 0.5*38.72+ -0.5*63.56+ -0.5*56.76= -8.4$


$\displaystyle SE(\hat{C_{2}}) = \sqrt{MSE* \sum_{i=1}^{g}\frac{c_{i}^{2}}{n_{i}}} = \sqrt{219 * (\frac{0.5^2}{25}} + \frac{0.5^2}{25}} + \frac{(-0.5)^2}{25}} + \frac{(-0.5)^2}{25})} = 2.9597$


CI: $-8.4 \pm 1.9799 \times 2.9597 = C_{2} \in (-14.26,-2.5401)$

From the confidence intervals, we can conclude that having 8 female companions prolongs a fly's lifespan and that having unreceptive companions negatively impacts a fly's lifespan.

### Q1c --- 4 points
$C =(\mu_{1p}+\mu_{1v})/2 - (\mu_{8p}+\mu_{8v})/2$ \dotfill (effect of receptive/not receptive across number of companions)


$t = \displaystyle \frac{\hat{C}}{\text{SE}(\hat{C})} = \frac{\sum_{i=1}^{g} c_{i}\bar{y}_{i\cdot}}{\sqrt{MSE* \sum_{i=1}^{g}\frac{c_{i}^{2}}{n_{i}}}}$


$\hat{C = \displaystyle \sum_{i=1}^{g} c_{i}\bar{y}_{i\cdot} = 0.5*\bar{y}_{1p\cdot} + 0.5*\bar{y}_{1v\cdot}+-0.5*\bar{y}_{8p\cdot} +-0.5*\bar{y}_{8v\cdot}= 0.5*63.56 + 0.5*64.80+ -0.5*56.76+ -0.5*38.72= 16.44$


$\displaystyle SE(\hat{C) = \sqrt{MSE* \sum_{i=1}^{g}\frac{c_{i}^{2}}{n_{i}}} = \sqrt{219 * (\frac{0.5^2}{25}} + \frac{0.5^2}{25}} + \frac{(-0.5)^2}{25}} + \frac{(-0.5)^2}{25})} = 2.9597$

$t = \displaystyle \frac{16.44}{2.9597}=5.5546$
```{r}
2*pt(5.5546,125-5,lower.tail = F)
```
We can conclude at the 95% confidence level that the effect of receptive vs nonreceptive companions on flies' lifespan does change with the number of companions since $p<0.05$.

### Q1d --- 4 points



## Question 2 --- 6 points

```{r}
ex3.5 = read.table("http://users.stat.umn.edu/~gary/book/fcdae.data/ex3.5", h=T)
ex3.5$trt = 15+15*ex3.5$trt 
```
Here is a plot of the data.
```{r, fig.width=3.5, fig.height=2.2, out.width='50%'}
ggplot(ex3.5, aes(x=trt, y=deg))+ geom_point() +
  ylab("Angle (Degree)")+xlab("Delay (minutes)")
```

