---
title: 'STAT 222 Spring 2022 HW4'
author: "Matthew Zhao"
fontsize: 12pt
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin=0.5in
fig_crop: no
---

```{r echo=F, include=F}
library(knitr)
library(ggplot2)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F, collapse=TRUE)
options(width=70, digits=6, scipen=8)
options(show.signif.stars = FALSE) # Show no stars for significance tests
```


# Question 1

## Q1a --- 4 points

Grand mean $\mu = \displaystyle \frac{\sum_{i=1}^{g} n_i \mu_i}{N} = \frac{45 * 3 + 33* 3 + 60* 4}{3 + 3 + 4} = 47.4$

```{r}
(45 * 3 + 33* 3 + 60* 4)/(3 + 3 + 4)
```


$\delta^2 = \displaystyle \frac{\sum_{i=1}^{g} n_i (\mu_i-\mu)^2 }{\sigma^2} = 35.4$
```{r}
((45-47.4)^2 * 3 + (33-47.4)^2* 3 + (60-47.4)^2* 4)/(36)
```


The power can be calculated using the R code below:
```{r}
pf(qf(1-0.01,3-1,10-3),3-1,10-3,ncp=35.4,lower.tail = F)
```


## Q1b --- 4 points

Grand mean $\mu = \displaystyle \frac{\sum_{i=1}^{g} n_i \mu_i}{N} = \frac{45 * n + 33* n + 60* n}{n + n + n} = \frac{138n}{3n} = 46$


$\delta^2 = \displaystyle \frac{\sum_{i=1}^{g} n_i (\mu_i-\mu)^2 }{\sigma^2} = \frac{n(45-46)^2 +n(33-46)^2 +n(60-46)^2}{36} = \frac{n+169n+196n}{36} = \frac{366n}{36} = 10.167n$
```{r}
# qf(alpha, g-1, N-g, lower.tail=F)
n=3
f_crit = qf(0.01, 3-1, 3*n-3, lower.tail=F)
pf(f_crit,3-1,3*n-3,ncp = 366*n/36,lower.tail=F)
```
```{r}
n=4
f_crit = qf(0.01, 3-1, 3*n-3, lower.tail=F)
pf(f_crit,3-1,3*n-3,ncp = 366*n/36,lower.tail=F)
```


We only reach a power of at least 0.95 with 4 subjects per group, meaning that we need to use at least 4 subjects per group to get to a power of 0.95.

# Question 2


```{r}
insulate = read.table("http://www.stat.uchicago.edu/~yibi/s222/insulate.txt", h=T)
```

## Q2a --- 4 points

```{r, message=FALSE, warning = F}
library(mosaic)
mean(failtime ~ as.factor(material), data=insulate)
anova(lm(failtime ~ as.factor(material), data=insulate))
```


$\displaystyle HSD_{\text{tukey}}=\frac{q_{\alpha}(g,N-g)}{\sqrt{2}} \times \sqrt{\text{MSE}(\frac{1}{n}+\frac{1}{n})}$
```{r}
(qtukey(1-0.1,5,20-5)/sqrt(2))*(sqrt(4167044*((1/4)+(1/4))))
insl_means <- insulate %>% group_by(material) %>% summarise(group_mean = mean(failtime))
for (x in 1:(nrow(insl_means)-1)) {
  for (y in (x+1):nrow(insl_means)) {
    print(paste0(as.character(insl_means[x,1]),"-",
                 as.character(insl_means[y,1]), " ",
                 round(abs(insl_means$group_mean[x]-
                             insl_means$group_mean[y]),3),
                 "   ",
                 ifelse(abs(insl_means$group_mean[x]-
                              insl_means$group_mean[y]) < 3906.62,
                        "Not Significant", "Significant")))
  }
}
```


Underline Diagram
```
     2      5       1         3         4
   6.25   10.75  159.75    2941.75   5723.00
  ---------------------------------
                          -------------------
```

## Q2b --- 2 points


```{r, warning=F}
library(MASS)
boxcox(failtime ~ as.factor(material), data=insulate)

```


Since our interval is centered around 0, we should use the log transformation for the response.

## Q2c --- 4 points

```{r, message=FALSE, warning = F}
mean(log10(failtime) ~ as.factor(material), data=insulate)
anova(lm(log10(failtime) ~ as.factor(material), data=insulate))
```
```{r}
(qtukey(1-0.1,5,20-5)/sqrt(2))*(sqrt(0.207*((1/4)+(1/4))))
insl_means <- insulate %>% mutate(failtime = log10(failtime)) %>% group_by(material) %>% summarise(group_mean = mean(failtime))
for (x in 1:(nrow(insl_means)-1)) {
  for (y in (x+1):nrow(insl_means)) {
    print(paste0(as.character(insl_means[x,1]),"-",
                 as.character(insl_means[y,1]), " ",
                 round(abs(insl_means$group_mean[x]-
                             insl_means$group_mean[y]),3),
                 "   ",
                 ifelse(abs(insl_means$group_mean[x]-
                              insl_means$group_mean[y]) < 0.870709,
                        "Not Significant", "Significant")))
  }
}
```


Underline Diagram
```
     2         5          1         3         4
  1.24245   1.90395    5.05159   7.71640   8.21401
  ------------------            -------------------
```

## Q2d --- 3 points
```{r}
ggplot(data = insulate, aes(x=as.factor(material), 
                            y=failtime)) + 
  geom_point() + 
  labs(x="material", y="failtime")

ggplot(data = insulate, aes(x=as.factor(material), 
                            y=log10(failtime))) + 
  geom_point() + 
  labs(x="material", y="log10(failtime)")

```


After transforming the response, the standard deviation across groups appears to be more similar than before, with the standard deviations of the groups fairly close at a glance (besides group 1). 

## Q2e --- 4 points

After the transformation, 1-2, 1-3, 1-5, 2-3, and 3-5 all become significant. All of the previously significant pairs are still significant. 
