---
title: "STAT 222 Spring 2022 HW7"
author: "Matthew Zhao"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin = .65in
fontsize: 12pt
---

```{r include=F}
library(tidyverse)
library(knitr)
library(MASS)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F, collapse=T)
options(scipen=6, digits=6)
# options(show.signif.stars = FALSE) # Show no stars for significance tests
```

```{r}
pr8.4 = read.table("http://users.stat.umn.edu/~gary/book/fcdae.data/pr8.4", header=T)
pr8.4$Time = factor(pr8.4$temp, labels = c("8hr","13hr"))
pr8.4$Airflow = factor(pr8.4$flow, labels = c("L","H"))
pr8.4$Laser = factor(pr8.4$laser, labels = c("old","new"))
```

## Q1 --- 5 points

```{r, message=FALSE}
library(mosaic)
mean(y ~ 1, data=pr8.4)
mean(y ~ Time, data=pr8.4)
mean(y ~ Airflow, data=pr8.4)
mean(y ~ Laser, data=pr8.4)
mean(y ~ Time+Airflow, data=pr8.4)
mean(y ~ Time+Laser, data=pr8.4)
mean(y ~ Airflow+Laser, data=pr8.4)
mean(y ~ Time+Airflow+Laser, data=pr8.4)
```

$\displaystyle \hat{\alpha}_{13hr} = \bar{y}_{13hr\cdot\cdot\cdot} - \bar{y}_{\cdot\cdot\cdot\cdot} = 0.19500-0.488125 = -0.293125$


$\displaystyle \hat{\beta\gamma}_{H,new} = \bar{y}_{\cdot H,new\cdot} - \bar{y}_{\cdot H\cdot\cdot}-\bar{y}_{\cdot\cdot new\cdot}+\bar{y}_{\cdot\cdot\cdot\cdot}=0.4425-0.47500- 0.47875+0.488125 = -0.023125$


$\displaystyle \hat{\alpha\beta\gamma}_{13hr,H,new} = \bar{y}_{13hr,H,new\cdot}-\bar{y}_{13hr,H\cdot\cdot}-\bar{y}_{13hr\cdot new\cdot} - \bar{y}_{\cdot H,new\cdot}+ \bar{y}_{13hr\cdot\cdot\cdot} + \bar{y}_{\cdot H\cdot\cdot} +\bar{y}_{\cdot\cdot new\cdot} - \bar{y}_{\cdot\cdot\cdot\cdot}= 0.120 -0.1725-  0.1925 - 0.4425 + 0.19500 +0.47500 + 0.47875 - 0.488125 = -0.026875$

## Q2 --- 5 points


```{r}
contrasts(pr8.4$Time) = contr.sum(2)
contrasts(pr8.4$Airflow) = contr.sum(2)
contrasts(pr8.4$Laser) = contr.sum(2)
lm1 = lm(y ~ Time*Airflow*Laser, data=pr8.4)
lm1$coef
options(digits=8) # increasing the number of digits in the output
anova(lm1)
```


$SS_{Time}=\displaystyle bcn \sum_{i=1}^{a} (\hat{\alpha}_{i})^{2} = bcn((\hat{\alpha}_{13hr})^{2} + (\hat{\alpha}_{13hr})^{2}) = 2 * 2 * 2 *((-0.293125)^2 + (0.293125)^2) = 1.374756$


$SS_{Airflow-Laser}=\displaystyle an \sum_{jk} (\hat{\beta\gamma}_{jk})^{2} = an((\hat{\beta\gamma}_{L,old})^{2} + (\hat{\beta\gamma}_{H,old})^{2} + (\hat{\beta\gamma}_{L,new} + (\hat{\beta\gamma}_{H,new})^{2})^{2}) = 2 * 2 *((-0.023125)^2 + (0.023125)^2 + (0.023125)^2 + (-0.023125)^2) = 0.00855625$


$SS_{Time-Airflow-Laser}=\displaystyle n \sum_{ijk} (\hat{\alpha\beta\gamma}_{ijk})^{2} = n((\hat{\alpha\beta\gamma}_{8hr,L,old})^{2} + (\hat{\alpha\beta\gamma}_{8hr,H,old})^{2} + (\hat{\alpha\beta\gamma}_{8hr,L,new})^{2} +  (\hat{\alpha\beta\gamma}_{8hr,H,new})^{2} + (\hat{\alpha\beta\gamma}_{13hr,L,old})^{2} + (\hat{\alpha\beta\gamma}_{13hr,H,old})^{2} + (\hat{\alpha\beta\gamma}_{13hr,L,new})^{2} +  (\hat{\alpha\beta\gamma}_{13hr,H,new})^{2}) = 2*(8*(0.026875)^2) = 0.01155625$

## Q3 --- 4 points

```{r}
lm1 = lm(y ~ Time*Airflow*Laser, data=pr8.4)
anova(lm1)
lmlog = lm(log(y) ~ Time*Airflow*Laser, data=pr8.4)
anova(lmlog)
```

```{r}
boxcox(lm1)
qqnorm(lm1$residuals)
qqline(lm1$residuals)
plot(lm1$fitted.values, lm1$residuals, xlab = "Fitted Values",
     ylab = "Residuals", main = "lm1")
```
```{r}
qqnorm(lmlog$residuals)
qqline(lmlog$residuals)
plot(lmlog$fitted.values, lmlog$residuals, xlab = "Fitted Values",
     ylab = "Residuals", main = "lmlog")
```


I believe that the lmlog ANOVA table is more trustworthy. This is because for the standard regression, the graph of fitted values against residuals reveals non-constant variance. Additionally, the qq plot shows that the data is light-tailed. The boxcox plot shows that 1 does lie within the 95% confidence interval; however, the interval is extremely wide and the other graphs provide evidence that the data violates our assumptions. Lastly, the fitted values vs residual plot shows constant variance, indicating that the ANOVA results should be more reliable than those of the base linear model. 

## Q4 --- 6 points

```{r, fig.width=5, fig.height=2.5, out.width='60%'}
par(mai=c(.6,.6,.01,.2),mgp=c(2,.6,0)) # for reducing the margin of the plot
with(pr8.4, interaction.plot(Time:Laser,Airflow,log(y), type="b"))
```


#### a)
Using a new vs old laser and changing the furnance airflow both do not result in a significant difference on delamination rate. This can be seen in the graph where there is virtually no change in log of delamination rate from old to new and by the fact that the lines for low and high airflow are on top of each other. There also does not appear to be an interaction between laser and airflow since the lines are nearly the same and parallel. 
#### b)
Yes there does appear to be an interaction since the lines are not parallel. This means that the effect of changing the laser on the log of delamination rate changes with different levels of airflow. 

#### c)
The three way interaction would be significant since the interaction between laser and airflow changes with different levels of firing profile time. Specifically, when firing profile time is 8hr, there does not appear to be any interaction between laser and furnance airflow, but when firing profile time is 13hr, there appears to be a significant interaction between laser and furnance airflow, supporting my initial statement. 

#### d)
Firing for 13hr instead of 8hr decreases delamination rate, cutting with a new laser results in a lower delamination rate compared with using an old laser, and having higher furnace airflow similarly lowers the delamination rate. Our conclusion is consistent.


## Q5 --- 5 points

```{r}
q <- qtukey(1-0.05,8,2*2*2*(2-1))
mse <- 0.02538
hsd <- q/sqrt(2) * sqrt(mse*((1/2) + (1/2)))
hsd
```

```{r}
pr8.4 %>% group_by(Time,Airflow,Laser) %>% summarise(.groups = 'keep',group_means = mean(log(y))) %>% arrange(desc(group_means))
```


13hr, high airflow, and new laser, along with 13hr, low airflow, and old laser, are significantly lower than the other log means. Since they are within the HSD of each other, they are not significantly different from each other. This means that we should make more replicates for these two so that we can find which combination is the best. 
