---
title: "STAT 222 Spring 2022 HW8"
author: "Matthew Zhao"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin = .65in
fontsize: 12pt
---

```{r include=F}
library(tidyverse)
library(knitr)
library(emmeans)
library(mosaic)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F, collapse=T)
options(scipen=6, digits=6)
# options(show.signif.stars = FALSE) # Show no stars for significance tests
```


```{r}
warpbreaks = read.table("http://www.stat.uchicago.edu/~yibi/s222/warpbreaks.txt", h=T)
warpbreaks$wool = as.factor(warpbreaks$wool)
warpbreaks$tension = factor(warpbreaks$tension, labels=c("L","M","H"))
lm1 = lm(breaks ~ wool*tension, data=warpbreaks)
```

## Q1 --- 5 points

We can use Tukey's HSD to control the FWER, where the HSD is given as $\displaystyle \frac{q_{g,dfE,\alpha}}{\sqrt{2}}\times\sqrt{\text{MSE}(\frac{1}{r}+\frac{1}{r})}$

```{r}
anova(lm1)
```


Calculating HSD:
```{r}
mse<-119.7
qval<-qtukey(1-0.02,6,48)/sqrt(2)
qval * sqrt(mse*(2/9))
```
```{r}
sort(mean(breaks~wool+tension,data=warpbreaks))
```


```
     2.H      1.M       1.H       2.L       2.M       1.L 
  18.7778   24.0000   24.5556   28.2222   28.7778   44.5556 
 -------------------------------------------------
                               -----------------------------
```

## Q2 --- 5 points

We can again use Tukey's HSD

```{r}
anova(lm1)
```


Calculating HSD:
```{r}
mse<-119.7
qval<-qtukey(1-0.02,3,48)/sqrt(2)
qval * sqrt(mse*(2/(9*2)))
```
```{r}
sort(mean(breaks~tension,data=warpbreaks))
```


```
      H         M        L 
  21.6667   26.3889   36.3889 
 -------------------
           -------------------
```


## Q3 --- 5 points

We use Bonferroni. 
```{r}
anova(lm1)
lm1emmean <- emmeans(lm1,~tension:wool)
summary(contrast(lm1emmean, method=list(Ls = c(1,0,0,-1,0,0),Ms = c(0,1,0,0,-1,0),Hs = c(0,0,1,0,0,-1)), adjust="bonf"),
infer=c(T,F), level=0.98)
```


$\mu_{1L} - \mu_{2L}$: (1.71,30.96)

$\mu_{1M} - \mu_{2M}$: (-19.40,9.85)

$\mu_{1H} - \mu_{2H}$: (-8.85,20.40)


We can conclude that there could potentially be an interaction since the CIs for the effect of changing wool type change with tension level. This means that there could be an interaction. However, since the CIs all overlap, with the first and third CIs having significant overlap, we cannot be sure that there is an interaction. 


## Q4 --- 6 points

```{r, fig.show='hold', fig.width=3.5, fig.height=2.5, out.width=c('40%','40%')}
par(mai=c(.6,.6,.1,.3),mgp=c(2,.6,0))
with(warpbreaks, interaction.plot(wool, tension, log(breaks), type="b"))
with(warpbreaks, interaction.plot(tension, wool, log(breaks), type="b"))
```

#### i)
$C_1\text{: } \sum_i c_{ij}$ 

$\text{j=L: } (1)\mu_{1L} + (-1)\mu_{2L} = 1 + (-1) = 0$

$\text{j=M: } (-1)\mu_{1M} + (1)\mu_{2M} = (-1) + 1 = 0$

$\sum_j c_{ij}$  

$\text{i=1: } (1)\mu_{1L} + (-1)\mu_{1M} = 1 + (-1) = 0$

$\text{i=2: } (-1)\mu_{2L} + (1)\mu_{2M} = (-1) + 1 = 0$


$C_2\text{: } \sum_i c_{ij}$ 

$\text{j=H: } (1)\mu_{1H} + (-1)\mu_{2H} = 1 + (-1) = 0$

$\text{j=M: } (-1)\mu_{1M} + (1)\mu_{2M} = (-1) + 1 = 0$

$\sum_j c_{ij}$  

$\text{i=1: } (1)\mu_{1H} + (-1)\mu_{1M} = 1 + (-1) = 0$

$\text{i=2: } (-1)\mu_{2H} + (1)\mu_{2M} = (-1) + 1 = 0$


$C_3\text{: } \sum_i c_{ij}$ 

$\text{j=L: } (0.5)\mu_{1L} + (-0.5)\mu_{2L} = 0.5 + (-0.5) = 0$

$\text{j=M: } (-1)\mu_{1M} + (1)\mu_{2M} = (-1) + 1 = 0$

$\text{j=H: } (-0.5)\mu_{1H} + (0.5)\mu_{2H} = (-0.5) + 0.5 = 0$

$\sum_j c_{ij}$  

$\text{i=1: } (0.5)\mu_{1L} + (0.5)\mu_{1H} (-1)\mu_{1M}= 0.5 + 0.5 + 1 = 0$

$\text{i=2: } (-0.5)\mu_{2L} + (-0.5)\mu_{2H} + (1)\mu_{2M}= (-0.5) + (-0.5) +1= 0$


#### ii)

We use Scheffe since we are performing contrasts after observing the data/based off of prior analysis. Scheffe's Adjustment can be given as $\hat{C} \pm \text{SE}(\hat{C})\times\sqrt{df_{\text{AB}}F_{df_{AB},dfE,\alpha}}$ where $\hat{C}= \sum_{ij}c_{ij}\bar{y}_{ij}$. 

```{r, warning=F}
anova(lm1)
mean(breaks~wool+tension,data=warpbreaks)
```


$\hat{C}_1 = 44.5556 - 28.2222 - (24.0000-28.7778) = 21.1112$


$\hat{C}_2 = 24.5556 - 18.7778 - (24.0000-28.7778) = 10.5556$


$\hat{C}_3 = \frac{\hat{C}_1 + \hat{C}_2}{2} = 15.8334$


$\text{SE}(\hat{C}) = \sqrt{\text{MSE}\times\sum_{ij}\frac{c_{ij}^2}{r}}$
```{r, warning=F}
mse <- 119.7
se_c12 <- sqrt(mse*(4*(1/9)))
se_c3 <- sqrt(mse*(((4*((0.5^2)/9))) + (2*(1/9))))
sprintf("C_1 t-stat: %f",21.1112/se_c12)
sprintf("C_2 t-stat: %f",10.5556/se_c12)
sprintf("C_3 t-stat: %f",15.8334/se_c3)

fcrit <- qf(0.04,(2-1)*(3-1),48,lower.tail = F)
scheffe <- sqrt((2-1)*(3-1)*fcrit)
sprintf("Scheffe Critical Value: %f",scheffe)
```


At FWER = 0.04, only the first contrast is significant. We can conclude that the change in effect of wool type on breaks is nonzero from low to medium. However, we cannot conclude that there is a significant interaction effect since only 1 of 3 contrasts testing for interaction were significant. 

## Q5 --- 1 point

The upper bound of the FWER for the entire analysis is no more than the sum of the FWER for each family so it is $0.02+0.02+0.02+0.04 = 0.1$.

## Q6 --- 3 points

Bonferroni. 

```{r, warning=F}
anova(lm1)
lm1emmean <- emmeans(lm1,~tension:wool)
summary(contrast(lm1emmean, method=list(C_1 = c(1,-1,0,-1,1,0), C_2 = c(0,-1,1,0,1,-1),C_3 = c(0.5,-1,0.5,-0.5,1,-0.5)), adjust="bonf"),
infer=c(F,T), level=0.96)
```


We can conclude at FWER=0.04 level that only the first contrast is significiant. This means that 