---
title: "STAT 222 Spring 2022 HW8"
author: "Matthew Zhao"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin = .65in
fontsize: 12pt
---

```{r include=F}
library(tidyverse)
library(knitr)
library(emmeans)
library(mosaic)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F, collapse=T)
options(scipen=6, digits=6)
# options(show.signif.stars = FALSE) # Show no stars for significance tests
```


```{r}
warpbreaks = read.table("http://www.stat.uchicago.edu/~yibi/s222/warpbreaks.txt", h=T)
warpbreaks$wool = as.factor(warpbreaks$wool)
warpbreaks$tension = factor(warpbreaks$tension, labels=c("L","M","H"))
lm1 = lm(breaks ~ wool*tension, data=warpbreaks)
```

## Q1 --- 5 points

We can use Tukey's HSD to control the FWER, where the HSD is given as $\displaystyle \frac{q_{g,dfE,\alpha}}{\sqrt{2}}\times\sqrt{\text{MSE}(\frac{1}{r}+\frac{1}{r})}$

```{r}
anova(lm1)
```


Calculating HSD:
```{r}
mse<-119.7
qval<-qtukey(1-0.02,6,48)/sqrt(2)
qval * sqrt(mse*(2/9))
```
```{r}
sort(mean(breaks~wool+tension,data=warpbreaks))
```


```
     2.H      1.M       1.H       2.L       2.M       1.L 
  18.7778   24.0000   24.5556   28.2222   28.7778   44.5556 
 -------------------------------------------------
                               -----------------------------
```

## Q2 --- 5 points

We can again use Tukey's HSD

```{r}
anova(lm1)
```


Calculating HSD:
```{r}
mse<-119.7
qval<-qtukey(1-0.02,3,48)/sqrt(2)
qval * sqrt(mse*(2/(9*2)))
```
```{r}
sort(mean(breaks~tension,data=warpbreaks))
```


```
      H         M        L 
  21.6667   26.3889   36.3889 
 -------------------
           -------------------
```


## Q3 --- 5 points

We again use Tukey.
```{r}
anova(lm1)
lm1emmean <- emmeans(lm1,~tension:wool)
summary(contrast(lm1emmean, method="pairwise", adjust="tukey"),
infer=c(T,F), level=0.98)
```


$\mu_{1L} - \mu_{2L}$: (-0.803,33.5)

$\mu_{1M} - \mu_{2M}$: (-21.914,12.4)

$\mu_{1H} - \mu_{2H}$: (-11.358,22.9)


We can conclude that there could be an interaction between wool type and level of tension since the CIs change with the level of tension. However, since the CIs all overlap, we cannot be sure that the interaction term is not insignificantly different than zero i.e. we cannot conclude that there is an interaction. 

## Q4 --- 6 points

```{r, fig.show='hold', fig.width=3.5, fig.height=2.5, out.width=c('40%','40%')}
par(mai=c(.6,.6,.1,.3),mgp=c(2,.6,0))
with(warpbreaks, interaction.plot(wool, tension, log(breaks), type="b"))
with(warpbreaks, interaction.plot(tension, wool, log(breaks), type="b"))
```

#### i)
$C_1\text{: } \sum_i c_{ij}$ 

$\text{j=L: } (1)\mu_{1L} + (-1)\mu_{2L} = 1 + (-1) = 0$

$\text{j=M: } (-1)\mu_{1M} + (1)\mu_{2M} = (-1) + 1 = 0$

$\sum_j c_{ij}$  

$\text{i=1: } (1)\mu_{1L} + (-1)\mu_{1M} = 1 + (-1) = 0$

$\text{i=2: } (-1)\mu_{2L} + (1)\mu_{2M} = (-1) + 1 = 0$


$C_2\text{: } \sum_i c_{ij}$ 

$\text{j=H: } (1)\mu_{1H} + (-1)\mu_{2H} = 1 + (-1) = 0$

$\text{j=M: } (-1)\mu_{1M} + (1)\mu_{2M} = (-1) + 1 = 0$

$\sum_j c_{ij}$  

$\text{i=1: } (1)\mu_{1H} + (-1)\mu_{1M} = 1 + (-1) = 0$

$\text{i=2: } (-1)\mu_{2H} + (1)\mu_{2M} = (-1) + 1 = 0$


$C_3\text{: } \sum_i c_{ij}$ 

$\text{j=L: } (0.5)\mu_{1L} + (-0.5)\mu_{2L} = 0.5 + (-0.5) = 0$

$\text{j=M: } (-1)\mu_{1M} + (1)\mu_{2M} = (-1) + 1 = 0$

$\text{j=H: } (-0.5)\mu_{1H} + (0.5)\mu_{2H} = (-0.5) + 0.5 = 0$

$\sum_j c_{ij}$  

$\text{i=1: } (0.5)\mu_{1L} + (0.5)\mu_{1H} (-1)\mu_{1M}= 0.5 + 0.5 + 1 = 0$

$\text{i=2: } (-0.5)\mu_{2L} + (-0.5)\mu_{2H} + (1)\mu_{2M}= (-0.5) + (-0.5) +1= 0$


#### ii)

We use Scheffe since we are performing contrasts after observing the data/based off of prior analysis. 

```{r, warning=F}
anova(lm1)
lm1emmean <- emmeans(lm1,~tension:wool)
summary(contrast(lm1emmean, method=list(C_1 = c(1,-1,0,-1,1,0), C_2 = c(0,-1,1,0,1,-1),C_3 = c(0.5,-1,0.5,-0.5,1,-0.5)), adjust="scheffe"),
infer=c(F,T), level=0.96)
```


At FWER = 0.04, only the first contrast is significant. We can conclude that the change in effect of wool type on breaks is nonzero from low to medium. 

## Q5 --- 1 point

The upper bound of the FWER for the entire analysis is no more than the sum of the FWER for each family so it is $0.02+0.02+0.02+0.04 = 0.1$.

## Q6 --- 3 points

Bonferroni. 

```{r, warning=F}
anova(lm1)
lm1emmean <- emmeans(lm1,~tension:wool)
summary(contrast(lm1emmean, method=list(C_1 = c(1,-1,0,-1,1,0), C_2 = c(0,-1,1,0,1,-1),C_3 = c(0.5,-1,0.5,-0.5,1,-0.5)), adjust="bonf"),
infer=c(F,T), level=0.96)
```

